---
title: "Food Finder"
author: "Olivia Hughes, Brooke Smith, Charlie Zhou"
date: "December 16, 2018"
output: html_document
---
#### Overview and Motivation:

As students in our first semester at the Harvard T.H. Chan School of Public Health, we wanted to investigate a topic specific to HSPH that was simple, while also intriguing and applicable to students. Before arriving at HSPH, we were aware of the free lunch offerings from their mention by current students in a prospective student video.  Now as enrolled students, we have first-hand knowledge of the numerous opportunities throughout the week to receive a free lunch. Our experiences attending events, and more specifically those with food, led to curiosity about which characteristics of an event may be predictive of the event offering a free meal.  Additionally, as many of us attending graduate school are financially disabled students, we thought it beneficial to create a resource with simple heuristics to use on any given day that predict the likelihood of an event providing a free lunch. 


#### Related Work/Inspiration:

A major benefit of being a Harvard student is having access to free lectures given by world class researchers, politicians, doctors, and professors.  At the Harvard T.H. Chan School of Public Health, a Student News email is sent each day to students detailing upcoming events.  As three new Harvard graduate students, we had heard the constant call to take advantage of these talks.  Not only would we be able to learn about new and exciting things happening in the world of public health, but perhaps we could get a free lunch as well.  Returning students notified us of this perk, as well as it being advertised in many events' description, and even in a video targeting prospective students.   


#### Initial Questions:

The question we are first and foremost trying to answer is where and when on campus can students find a free lunch. To do this, we planned to collect information only regarding events accompanied by free meals. Due to data size constraints, however, our question broadened to encompass an analysis of all events on campus, regardless of food offerings.  As such, the final data incorporates levels of free food available, that is, a full meal option versus simply refreshments or other snacks, which we collectively refer to as "snacks".  Subsequently, we explored a variety of other event characteristics. Are they run by student groups or the Harvard institution? Are they on campus or off campus? What time of day are they held? All of this culminated in analysis of the best opportunities to find free food as well as a harder look at the details of the plethora of events hosted by HSPH and beyond. 


#### Data:

Our source of data for this project was the daily Student News emails. Each group member was responsible for collecting one third of the data for all the events starting from the first email of the semester until November 30th, 2018. Data collection was broken down into 22 different variables describing location, topic and/or speaker, and additional offerings (e.g. workshop, free lunch, refreshments, etc.).  

```{r} 

#reading in data 

library(readr) 

masterdata <- read_csv("https://raw.githubusercontent.com/OliviaMoriah/Food-Finder/master/BST%20260%20Group%20Project%20-%20Data.csv")

``` 

Once we began working with the data, we noticed errors that would affect analyses i.e. a misspelled weekday, missing event end time, and host information so we performed some data cleaning work to amend for this.. 

```{r echo=FALSE} 
library(dplyr)
```

```{r}
#Check Weekday 

unique(masterdata$Weekday) 

masterdata <- masterdata%>% mutate(Weekday = recode(Weekday, `Thurday`="Thursday")) 

#duration = duration of event, if na then lasts 1 hour 

masterdata <- masterdata %>% mutate(duration=as.numeric(TimeEnd-TimeStart)/60) %>% mutate(duration=ifelse(is.na(duration), 60, duration)) 

```

After data collection and upon the start of our analysis, we realized many variables we collected would be useless in a coding environment so we decided to make new binary variables from existing two-level variables, and changed variables with multiple levels into factors. 

```{r}

## Making Institution and Student binary and coding blanks as "Institution"  

masterdata <- masterdata %>% mutate(instituteyn=ifelse(InstituteVStudent=="Institution"|InstituteVStudent=="institution"|is.na(InstituteVStudent), 1, 0), studentyn=ifelse(InstituteVStudent=="Student"|InstituteVStudent=="student",1,0)) 

#turn food into binary 

masterdata <- masterdata%>%mutate(foodyn=ifelse(Food=="Y"|Food=="y", 1,0))%>% mutate(foodyn=ifelse(!is.na(foodyn), foodyn, 0)) 

#turn snack into binary 

masterdata<- masterdata%>%mutate(snack=ifelse(MealSnake=="snack", 1,0))%>% mutate(snack=ifelse(!is.na(snack), snack, 0)) 

#turn meal into binary 

masterdata<- masterdata%>%mutate(meal=ifelse(MealSnake=="meal", 1,0))%>% mutate(meal=ifelse(!is.na(meal), meal, 0)) 

#turn Workshop into binary 

masterdata <- masterdata%>%mutate(workshopyn=ifelse(Workshop=="Y"|Workshop=="y", 1, 0)) 

#turn Series into binary 

masterdata <- masterdata%>%mutate(seriesyn=ifelse(Series=="Y"|Series=="y", 1, 0)) 
#turn Open into binary 

masterdata <- masterdata%>%mutate(openyn=ifelse(Open=="Y"|Open=="y", 1, 0)) 

#turn RSVP into binary 

masterdata <- masterdata%>%mutate(RSVPyn=ifelse(RSVP=="Y"|RSVP=="y", 1, 0)) 

#turn Speaker into binary 

masterdata <- masterdata%>%mutate(speakeryn=ifelse(Speaker=="Panel"|Speaker=="panel"|is.na(Speaker)==TRUE, 0, 1)) 

#Creating condensed topic column: FourTopics and making Weekday into factors 

levels(as.factor(masterdata$FourTopics)) 

masterdata$FourTopics <- as.factor(masterdata$FourTopics) 

masterdata$Weekday <- as.factor(masterdata$Weekday) 

levels(masterdata$Weekday) 

#reorder Weekdays 

masterdata$Weekday=factor(masterdata$Weekday, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")) 
```

We felt that week of the semester would be an interesting variable to include, so created a variable based on the existing date variable. 

```{r} 

#adding Date as week of the semester 

range(masterdata$Date) 

masterdata <- masterdata %>% mutate(SemesterWeek = ifelse(Date<"2018-12-16", ifelse(Date<"2018-12-09", ifelse(Date<"2018-12-02", ifelse(Date < "2018-11-25", ifelse(Date<"2018-11-18", ifelse(Date<"2018-11-11", ifelse(Date<"2018-11-04", ifelse(Date<"2018-10-28", ifelse(Date<"2018-10-21", ifelse(Date<"2018-10-14", ifelse(Date<"2018-10-07", ifelse(Date<"2018-09-30", ifelse(Date<"2018-09-23", ifelse(Date<"2018-09-16", ifelse(Date<"2018-09-09", ifelse(Date<"2018-09-02", 0, 1), 2), 3), 4), 5), 6), 7), 8), 9), 10), 11), 12), 13), 14), 15), 16)) 
```
 

#### Exploratory Analysis: 

In our intial analysis, we made use of methods for observing our descriptive statistics using histograms and simple proportions. Below a multitude of statistics are explored using our original data. Some we felt were not as compelling or suffered from a dearth of data, i.e. in the list of speakers at events. Once we found the statistics that we felt would be the most useful, relevant, and interesting to food event prediction for students, we used more sophisticated methods for evaluating this data using RShiny and ggplot2.


```{r} 
library(ggplot2)

##Underlying question: WHICH LEVELS OF EACH VARIABLE HAVE MORE FOOD? 

#descriptive stats regarding Speaker-about 50% of events with a speaker have food

foodprop_speaker <- masterdata %>% group_by(speakeryn) %>% summarise(propfoodevents = sum(foodyn==1)/n()) 

#descriptive stats regarding Workshop - about 50% of events considered a workshop have food 

foodprop_workshop <- masterdata %>% group_by(workshopyn) %>% summarise(propfoodevents = sum(foodyn==1)/n()) 

#descriptive stats regarding Topic (in 5 categories) 

numtopic <- masterdata %>% group_by(FourTopics) %>% summarise(number_of_events = n()) 

proptopic <- masterdata %>% group_by(FourTopics) %>% summarise(number_of_events = n()/270) 

#about half of all events are "academic"

foodprop_topic <- masterdata %>% group_by(FourTopics) %>% summarise(propfoodevents = sum(foodyn==1)/n()) 

#Just under half of academic and welfare events have food 

#Just under half of social and career events have food 

#all religious events have food (although there are only 7 events in this category)

masterdata %>% filter(foodyn==1) %>% ggplot() + geom_bar(aes(FourTopics)) 

#yes, academic has the most food events, but it also has the most events in general 

``` 

```{r} 

### LOOKING AT WEEK OF THE SEMESTER 

#proportion of events by week 

masterdata %>% group_by(SemesterWeek) %>% summarise(propev = n()) 

#the 2nd to middle week of the semester has the most events 

#proportion of food events by week 

masterdata %>% group_by(SemesterWeek) %>% summarise(propfood = sum(foodyn==1)/n()) 

#from the 5th week on, over half of a week's events provide some sort of food 

#proportion of meal events by week 

masterdata %>% group_by(SemesterWeek) %>% summarise(propmeal = sum(meal==1)/n()) 

#the 9th week of the semester (1st week of Fall 2), is the week with the highest proportion of meal events 

masterdata %>% group_by(SemesterWeek) %>% summarise(nmeal = sum(meal==1)) 

#weeks 5-9 have the highest number of meals offered from 7 to 9 meals per week 

``` 


```{r} 

#creating a data frame for rshiny plots looking at the number of food events on a weekday in a specific week 

numfood_by_semweek <- masterdata %>% filter(foodyn==1) %>% group_by(SemesterWeek) %>% summarise(Monday = sum(Weekday=="Monday"), Tuesday = sum(Weekday=="Tuesday"), Wednesday = sum(Weekday=="Wednesday"), Thursday = sum(Weekday=="Thursday"), Friday = sum(Weekday=="Friday"), Saturday = sum(Weekday=="Saturday")) 

#flipping rows and columns 

numfood_by_semweek_flip <- data.frame(t(numfood_by_semweek[-1])) 

#renaming Columns (note: week 0, 12, 15 had no food events) 

colnames(numfood_by_semweek_flip) <- c("Week 1","Week 2","Week 3","Week 4","Week 5","Week 6","Week 7","Week 8","Week 9","Week 10","Week 11","Week 13","Week 14") 

rows <- rownames(numfood_by_semweek_flip) 

rows <- factor(rows, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")) 

``` 

```{r echo= FALSE} 

library(shiny) 

library(shinyWidgets) 

library(ggplot2) 
```
```{r}

ui <- fluidPage(     
  

  # Give the page a title 

  titlePanel("Food Events by Day of the Week and Week of the Semester"), 

   

  # Generate a row with a sidebar 

  sidebarLayout(       

     

    # Define the sidebar with one input 

    sidebarPanel( 

      sliderTextInput("week", "Week:", choices=colnames(numfood_by_semweek_flip), animate = TRUE), 

      hr(), 

      helpText("Note: Weeks 0, 12, 15 had no events")), 

     

    # Create a spot for the barplot 

    mainPanel( 

      plotOutput("propPlot")) 

  ) 

) 

 

server <- # Define a server for the Shiny app 

  function(input, output) { 

     

    # Fill in the spot we created for a plot 

    output$propPlot <- renderPlot({ 

       

      # Render a barplot 

      ggplot() + geom_col(aes(x = rows, y = numfood_by_semweek_flip[,input$week])) + xlab("") + ylab("Number of Food Events") + scale_y_continuous(limits = c(0, 10)) 

    }) 

  } 

 
shinyApp(ui=ui,server=server) 

``` 


```{r echo=FALSE}     

library(tidyverse)     

library(readr)     

library(readxl)     

library(lubridate)     

library(dplyr)     

library(ggplot2) 

library(gridExtra)

```   
 

```{r}     

#Bulk exploratory analysis

a<-masterdata%>% summarize(total_events=n(), events_with_food=sum(foodyn), events_with_snack=sum(snack), events_with_meal=sum(meal), proportion_total=n()/n(), proportion_with_food=sum(foodyn)/n(), proportion_with_snack=sum(snack)/n(), proportion_with_meal=sum(meal)/n()) 

##Total number analysis 

total_number<-a %>% select(-proportion_total, -proportion_with_food, -proportion_with_meal, -proportion_with_snack) %>% gather(variable, value, total_events, events_with_food, events_with_snack, events_with_meal) 

total_number$variable=factor(total_number$variable, levels=c("total_events", "events_with_food", "events_with_meal", "events_with_snack")) 

total_number<-total_number%>%mutate(variable = recode(variable, `total_events`="Total", `events_with_food` = "Food", `events_with_meal`="Meal", `events_with_snack`="Snack"))     

##Proportion analysis 

total_proportion<- a %>% select(-total_events, -events_with_food, -events_with_snack, -events_with_meal) %>% gather(variable, value, proportion_total, proportion_with_food, proportion_with_meal, proportion_with_snack) 

total_proportion$variable=factor(total_proportion$variable, levels=c("proportion_total", "proportion_with_food", "proportion_with_meal", "proportion_with_snack")) 

total_proportion<-total_proportion%>%mutate(variable = recode(variable, `proportion_total`="Total", `proportion_with_food` = "Food", `proportion_with_meal`="Meal", `proportion_with_snack`="Snack")) 

``` 


```{r} 

library(ggthemes)

#Bulk Analysis Final 

a2<-masterdata%>% summarize(total_events=n(), events_with_food=sum(foodyn), events_with_snack=sum(snack), events_with_meal=sum(meal), no_food=n()-sum(foodyn), proportion_total=n()/n(), proportion_with_food=sum(foodyn)/n(), proportion_with_snack=sum(snack)/n(), proportion_with_meal=sum(meal)/n(), proportion_no_food=(n()-sum(foodyn))/n()) 

 total_proportion2<- a2 %>% select(-total_events, -events_with_food, -events_with_snack, -events_with_meal, no_food) %>% gather(variable, value, proportion_total, proportion_with_food, proportion_with_meal, proportion_with_snack, proportion_no_food) 

total_proportion2$variable=factor(total_proportion2$variable, levels=c("proportion_total", "proportion_with_food", "proportion_with_meal", "proportion_with_snack", "proportion_no_food")) 

total_proportion2<-total_proportion2%>%mutate(variable = recode(variable, `proportion_total`="Total", `proportion_with_food` = "Food", `proportion_with_meal`="Meal", `proportion_with_snack`="Snack", `proportion_no_food`="No Food")) 


# Bar chart depicting the proportion of events with food   

total_proportion2 %>% filter(variable!="Total"& variable!="Food") %>% arrange(value) %>% ggplot(aes(variable, value,fill=variable)) + 

  geom_bar(stat="identity", alpha=0.75) + 

  theme(legend.position="none") + 

  theme(axis.text.x=element_text(face="bold", size=12), axis.text.y=element_text(face="bold", size=12))+ 

   scale_color_economist()+
 scale_fill_manual(values=c("greenyellow", "seagreen2", "firebrick3"))+

  xlab ("")+ 

  ylab ("")+ 

  ggtitle("Events and Food Status by Proportion") 

``` 

The above graph represents the proportion of events without food, with meal, or with snack.

```{r} 

#Weekday analysis exploratory  

masterdata$Weekday=factor(masterdata$Weekday, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")) 

weekday_analysis<-masterdata %>% group_by(Weekday) %>% summarize(events_on_day=n(), food_on_day=sum(foodyn), proportion_with_food=mean(foodyn), snack_on_day=sum(snack), proportion_with_snack=mean(snack), meal_on_day=sum(meal), proportion_with_meal=mean(meal))     


weekday_analysis<-weekday_analysis %>% select(-events_on_day, -food_on_day, -snack_on_day, -meal_on_day) %>% filter(Weekday== c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))%>% gather(Variable, Value, proportion_with_food, proportion_with_snack, proportion_with_meal) %>% arrange(Weekday)

  
levels(weekday_analysis$Weekday)

weekday_analysis %>% ggplot(aes(Weekday, Value, fill=Variable))+ 

  geom_bar(position="dodge", stat="identity", alpha=0.75)+ 

  theme(axis.text.x=element_text(face="bold", size=12), axis.text.y=element_text(face="bold", size=12))+ 

  scale_fill_manual(values=c("cornflowerblue", "greenyellow", "seagreen2"), name="", labels=c("Food", "Meal", "Snack"))+ 

  xlab ("")+ 

  ylab ("")+ 

  ggtitle("Probability of Food by Weekday") 



``` 

Exploratory graph of the proportion of events with food, with meals, and with snacks by day of the week.

```{r} 

#Weekday analysis final  

weekday_analysis2<-masterdata %>% group_by(Weekday) %>% summarize(total_events=n(), events_with_food=sum(foodyn), events_with_snack=sum(snack), events_with_meal=sum(meal), no_food=n()-sum(foodyn), proportion_total=n()/n(), proportion_with_food=sum(foodyn)/n(), proportion_with_snack=sum(snack)/n(), proportion_with_meal=sum(meal)/n(), proportion_no_food=(n()-sum(foodyn))/n()) 


weekday_analysis2a<-weekday_analysis2 %>% select(-total_events, -events_with_food, -events_with_snack, -events_with_meal, -no_food) %>% filter(Weekday== c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))%>% gather(Variable, Value, proportion_total, proportion_with_food, proportion_with_snack, proportion_with_meal, proportion_no_food) %>% arrange(Weekday) 

  
weekday_analysis2a$Variable=factor(weekday_analysis2a$Variable, levels=c("proportion_total", "proportion_with_food", "proportion_with_meal", "proportion_with_snack", "proportion_no_food")) 

  
weekday_analysis2a %>% filter(Variable!="proportion_total" & Variable!="proportion_with_food") %>% ggplot(aes(Weekday, Value, fill=Variable))+ 

  geom_bar(position="dodge", stat="identity", alpha=0.75)+ 

  theme(axis.text.x=element_text(face="bold", size=12), axis.text.y=element_text(face="bold", size=12))+ 

  scale_fill_manual(values=c("greenyellow", "seagreen2", "firebrick3"),name="", labels=c("Meal", "Snack", "No food"))+ 

  xlab ("")+ 

  ylab ("")+ 

  ggtitle("Probability of Food by Weekday") 

``` 

Depiction of the proportion of events without food, with meal and with snacks by day of the week.  

```{r} 

#Weekday number analysis final 

weekday_analysis2 

weekday_analysis2b<-weekday_analysis2 %>% select(Weekday, events_with_snack, events_with_meal, no_food) %>% filter(Weekday== c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))%>%gather(Variable, Value, events_with_snack, events_with_meal, no_food) %>% arrange(Weekday) 

weekday_analysis2b$Variable=factor(weekday_analysis2b$Variable, levels=c("events_with_snack", "events_with_meal", "no_food")) 

 weekday_analysis2b %>% ggplot(aes(Weekday, Value, fill=Variable))+ 

  geom_bar(position="dodge", stat="identity", alpha=0.75)+ 

  theme(axis.text.x=element_text(face="bold", size=12), axis.text.y=element_text(face="bold", size=12))+ 

  scale_fill_manual(values=c("greenyellow", "seagreen2", "firebrick3"), name="", labels=c("Meal", "Snack", "No food"))+ 

  xlab ("")+ 

  ylab ("")+ 

  ggtitle("Events by Weekday") 

``` 

Number of events without food, with snacks and with meals by day of the week. 

```{r} 

#Start Time Analysis by Day 


masterdata<-masterdata%>%mutate(foodynfactor=as.factor(foodyn)) 

masterdata %>% ggplot(aes(as.numeric(TimeStart), fill=foodynfactor))+ 

  geom_density(alpha=0.5, adjust=1, position="stack")+ 

  scale_x_continuous(limits=c(as.numeric(hms("06:00:00")),as.numeric(hms("22:00:00"))), breaks=seq(21600, 79200, 3600), labels= c("6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "12PM", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM"))+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"), name="Food Status", labels=c("No", "Yes")) + 

  scale_y_continuous(labels = NULL) + 

  theme(panel.background = element_blank(), panel.grid.major.y = element_blank())+ 

  xlab("")+ 

  ylab("")+ 

  ggtitle("All") 

``` 

Probability distribution of events and of food at those events by start time of the event across all days.

```{r} 

masterdata%>% filter(Weekday=="Monday") %>% ggplot(aes(as.numeric(TimeStart), fill=foodynfactor))+ 

  geom_density(alpha=0.5, adjust=1, position="stack")+ 

  scale_x_continuous(limits=c(as.numeric(hms("06:00:00")),as.numeric(hms("22:00:00"))), breaks=seq(21600, 79200, 3600), labels= c("6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "12PM", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM"))+ 

  theme(axis.text.x=element_text(size=5))+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"),name="Food Status", labels=c("No", "Yes")) + 

  scale_y_continuous(labels = NULL) + 

  theme(panel.background = element_blank(), panel.grid.major.y = element_blank())+ scale_color_economist()+ 

  xlab("")+ 

  ylab("")+ 

  ggtitle("Monday") 

  

ggsave("Monday.png", scale=0.7, dpi=300) 

  

masterdata%>% filter(Weekday=="Tuesday") %>% ggplot(aes(as.numeric(TimeStart), fill=foodynfactor))+ 

  geom_density(alpha=0.5, adjust=1, position="stack")+ 

  scale_x_continuous(limits=c(as.numeric(hms("06:00:00")),as.numeric(hms("22:00:00"))), breaks=seq(21600, 79200, 3600), labels= c("6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "12PM", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM"))+ 

  theme(axis.text.x=element_text(size=5))+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"),name="Food Status", labels=c("No", "Yes")) + 

  scale_y_continuous(labels = NULL) + 

  theme(panel.background = element_blank(), panel.grid.major.y = element_blank())+ scale_color_economist()+ 

  xlab("")+ 

  ylab("")+ 

  ggtitle("Tuesday") 

  

ggsave("Tuesday.png", scale=0.7, dpi=300) 

  

masterdata%>% filter(Weekday=="Wednesday") %>% ggplot(aes(as.numeric(TimeStart), fill=foodynfactor))+ 

  geom_density(alpha=0.5, adjust=1, position="stack")+ 

  scale_x_continuous(limits=c(as.numeric(hms("06:00:00")),as.numeric(hms("22:00:00"))), breaks=seq(21600, 79200, 3600), labels= c("6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "12PM", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM"))+ 

  theme(axis.text.x=element_text(size=5))+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"),name="Food Status", labels=c("No", "Yes")) + 

  scale_y_continuous(labels = NULL) + 

  theme(panel.background = element_blank(), panel.grid.major.y = element_blank())+ scale_color_economist()+ 

  xlab("")+ 

  ylab("")+ 

  ggtitle("Wednesday") 

  

ggsave("Wednesday.png", scale=0.7, dpi=300) 

  

masterdata%>% filter(Weekday=="Thursday") %>% ggplot(aes(as.numeric(TimeStart), fill=foodynfactor))+ 

  geom_density(alpha=0.5, adjust=1, position="stack")+ 

  scale_x_continuous(limits=c(as.numeric(hms("06:00:00")),as.numeric(hms("22:00:00"))), breaks=seq(21600, 79200, 3600), labels= c("6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "12PM", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM"))+ 

  theme(axis.text.x=element_text(size=5))+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"),name="Food Status", labels=c("No", "Yes")) + 

  scale_y_continuous(labels = NULL) + 

  theme(panel.background = element_blank(), panel.grid.major.y = element_blank())+ scale_color_economist()+ 

  xlab("")+ 

  ylab("")+ 

  ggtitle("Thursday") 

  

ggsave("Thursday.png", scale=0.7, dpi=300) 

            

masterdata%>% filter(Weekday=="Friday") %>% ggplot(aes(as.numeric(TimeStart), fill=foodynfactor))+ 

  geom_density(alpha=0.5, adjust=1, position="stack")+ 

  scale_x_continuous(limits=c(as.numeric(hms("06:00:00")),as.numeric(hms("22:00:00"))), breaks=seq(21600, 79200, 3600), labels= c("6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "12PM", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM"))+ 

  theme(axis.text.x=element_text(size=5))+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"),name="Food Status", labels=c("No", "Yes")) + 

  scale_y_continuous(labels = NULL) + scale_color_economist()+ 

  theme(panel.background = element_blank(), panel.grid.major.y = element_blank())+ 

  xlab("")+ 

  ylab("")+ 

  ggtitle("Friday") 

  

ggsave("Friday.png", scale=0.7, dpi=300) 

  

masterdata%>% filter(Weekday=="Saturday") %>% ggplot(aes(as.numeric(TimeStart), fill=foodynfactor))+ 

  geom_density(alpha=0.5, adjust=1, position="stack")+ 

  scale_x_continuous(limits=c(as.numeric(hms("06:00:00")),as.numeric(hms("22:00:00"))), breaks=seq(21600, 79200, 3600), labels= c("6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "12PM", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM"))+ 

  theme(axis.text.x=element_text(size=5))+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"),name="Food Status", labels=c("No", "Yes")) + 

  scale_y_continuous(labels = NULL) + 

  theme(panel.background = element_blank(), panel.grid.major.y = element_blank())+ scale_color_economist()+ 

  xlab("")+ 

  ylab("")+ 

  ggtitle("Saturday") 

  

ggsave("Saturday.png", scale=0.7, dpi=300) 

``` 

Probability distribution of events and of food at those events by start time of the event by weekday. This will produce graphs that will be used for an R shiny animation.

```{r}     
#needed for rshiny
path <-getwd()
```  

```{r} 

#R shiny 

library(shiny) 

library(shinyWidgets) 

  

ui <- fluidPage( 

  sliderTextInput("Weekday", "Day of the Week:", choices=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),selected = c("Monday"), animate = TRUE, grid = TRUE, hide_min_max = TRUE, from_fixed = FALSE,to_fixed = FALSE, from_min = NULL, from_max = NULL, to_min = NULL,to_max = NULL, force_edges = FALSE, width = 1000, pre = NULL,post = NULL, dragRange = TRUE), 

  plotOutput("graph")) 

  

values <- reactiveValues("Monday" = "Monday.png", "Tuesday"="Tuesday.png", "Wednesday"="Wednesday.png", "Thursday"="Thursday.png", "Friday" = "Friday.png", "Saturday"="Saturday.png") 

  

paste('red', 'dead', sep='') 

paste('weekday', '.png', sep='') 

  

  

server <- function(input, output) { 

  output$graph <- renderImage ({ 

    list(src= 

    file.path(path, paste(input$Weekday, '.png', sep=''))) 

  }, deleteFile=FALSE) 

} 

  

shinyApp(ui=ui,server=server) 

  

``` 

R shiny animation/interactive slider of probability distribution graphs of events and food at those events.

Takeaway - Bimodal peak - lunchtime and late afternoon/evening 

```{r} 
#Snack vs Meal over Time 

masterdata<-masterdata%>%mutate(mealfactor=as.factor(meal)) 

 
masterdata %>% filter(foodyn==1) %>% ggplot(aes(as.numeric(TimeStart), fill=mealfactor))+ 

  geom_density(alpha=0.5, adjust=1)+ 

  scale_x_continuous(limits=c(as.numeric(hms("06:00:00")),as.numeric(hms("22:00:00"))), breaks=seq(21600, 79200, 3600), labels= c("6AM", "7AM", "8AM", "9AM", "10AM", "11AM", "12PM", "1PM", "2PM", "3PM", "4PM", "5PM", "6PM", "7PM", "8PM", "9PM", "10PM"))+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"), name="Food Type", labels=c("Snack", "Meal")) + 

  scale_y_continuous(labels = NULL) + 

  theme(panel.background = element_blank(), panel.grid.major.y = element_blank())+ scale_color_economist()+ 

  xlab("")+ 

  ylab("")+ 

  ggtitle("Food Type by Time of Day") 

``` 

Distribution of meals and snacks by start time of the event.

Takeaway - meals happen at mealtimes, and snacks happen throughout afternoon - pretty stable 

```{r} 

#By calendar date 

calendar_date<-masterdata%>%group_by(Date) %>% summarize(events=n(), food=sum(foodyn)) 

all_dates_in_range <- data.frame (Date= seq.Date(min(calendar_date$Date), max(calendar_date$Date), by="day")) 

calendar_date<-left_join (all_dates_in_range, calendar_date, by="Date") 

calendar_date<-calendar_date%>%mutate(events=ifelse(is.na(events), 0, events), food=ifelse(is.na(food), 0, food)) 

calendar_date<-calendar_date%>%gather(variable, value, events, food)%>%arrange(Date) 

  

sundays<-seq(as.Date("2018/8/26"), as.Date("2018/12/12"), by = "week") 

sundays<-as.data.frame(sundays) 

  

calendar_date%>%ggplot(aes(Date, value, fill=variable)) +  

  geom_area(alpha=0.5)+ 

  geom_vline(data=sundays, aes(xintercept=sundays), linetype="dashed")+ 

  scale_fill_manual(values=c("orangered2","cornflowerblue"),name="", labels=c("Events", "Food")) + 
scale_color_economist()+ 
  ylab("")+ 

  xlab("")+ 

  ggtitle("Events by Calendar Date") 

  

```   

Number of events and events with food by calendar date.

No events over Thanksgiving or on Sundays (denoted by dotted lines).

```{r} 

## Who more likely to host event? - Proportion of events school vs. institution 

## Institution more likely to host with 72.5% of events 

inst_v_school <- masterdata %>% summarize(Institute=sum(instituteyn, na.rm=T)/n(), Student=sum(studentyn, na.rm=T)/n()) 

##Whose events more likely to have food? 
## Institute with 66% 

inst_v_school_food <- masterdata %>% filter(foodyn==1) %>% summarize(Inst_food_prop = sum(instituteyn, na.rm = T)/n(), Student_food_prop=sum(studentyn, na.rm = T)/n()) 

``` 


```{r warning=FALSE} 

library(stringr) 

## Converting location to a vector to allow for string processing 

newvec <- c(masterdata$Location) 

## Data frame specific to locations 
## Using string processing to pull out building location of the event based on Location listed in our spreadsheet 

event_df <- masterdata %>% select(ID, Location, foodyn, snack,meal) %>%mutate(Kresge = ifelse(str_detect(newvec, "Kresge"),1,0), Countway = ifelse(str_detect(newvec, "Countway"),1,0), FXB=ifelse(str_detect(newvec,"FXB"),1,0), Leadership = ifelse(str_detect(newvec, "Leadership"),1,0), ODI=ifelse(str_detect(newvec,"ODI"),1,0)) 

##Kresge total events - 121 events 

sum(event_df$Kresge, na.rm = T) 

## Total HSPH campus events 

sum(event_df$Kresge + event_df$Countway + event_df$FXB + event_df$Leadership + event_df$ODI, na.rm = TRUE) 


## adding off-campus  and "other"" event column 

event_df <- event_df %>% mutate(PHOffcampus = ifelse((Kresge==1 | Countway==1 | ODI==1 | Leadership==1 | FXB==1), 0, 1), OtherHSPH = ifelse((ODI==1 | Leadership==1), 1, 0))  

## Condensing events to format readable for ggplot 

event_df_condensed <- event_df %>% select(Kresge, Countway, FXB, PHOffcampus, OtherHSPH) %>% gather(Building, value, Kresge, Countway, OtherHSPH, FXB, PHOffcampus) 

## Bar chart of all event locations 

BC_EL <-event_df_condensed %>% arrange(value) %>% ggplot(aes(Building, value,fill=Building)) + 

geom_bar(stat="identity") + 

theme(legend.position="none") + 
  theme(axis.text.x=element_text(face="bold", size=12), axis.text.y=element_text(face="bold", size=12))+ 

  scale_fill_economist()+

  xlab("")+ 

ylab ("Number of Events")+ 

  scale_x_discrete(labels=c("Countway", "FXB", "Kresge", "Other HSPH", "Off-campus"))+ scale_color_economist()+ 

ggtitle("Event Location Distribution") 


## Number of events at each location with food                                  

food_events <- event_df %>% mutate(KresgeFood=ifelse(Kresge==1 & foodyn==1, 1, 0), OtherFood=ifelse(ODI==1 | Leadership==1 & foodyn==1, 1, 0), CountwayFood=ifelse(Countway==1 & foodyn==1,1,0), FXBFood=ifelse(FXB==1 & foodyn==1,1,0), PHOffcampusFood=ifelse(PHOffcampus==1&foodyn==1,1,0)) 


## Condensing events to format readable for ggplot 

food_events_condensed <- food_events %>% select(KresgeFood, OtherFood,CountwayFood, FXBFood, PHOffcampusFood) %>% gather(Building, value, KresgeFood, OtherFood,CountwayFood, FXBFood, PHOffcampusFood) 
 

## Bar chart of food event locations 

BC_FD <-food_events_condensed %>% arrange(value) %>% ggplot(aes(x=Building,y=value, fill=Building)) + 

geom_histogram(stat="identity") + 

theme(legend.position="none") + 

theme(axis.text.x=element_text(face="bold", size=12), axis.text.y=element_text(face="bold", size=12))+ 
    scale_fill_economist()+
  xlab("")+ 

ylab ("Number of Events")+ 

scale_x_discrete(labels=c("Countway", "FXB", "Kresge", "Other HSPH", "Off-campus")) + scale_color_economist()+ ylim(0, 115) +

ggtitle("Event with Food Location Distribution") 

grid.arrange(BC_EL, BC_FD)

``` 

     

#### Final Analysis:

Major takeaways from our data analysis are as follows:
Our final analysis showed us that approximately 50% of events offered free food. Of these events, unsurprisingly, your best chance of getting a full meal at HSPH is during the lunch period from 12:00pm to 2:00pm and the dinner period from 5:00 to 8:00pm. Peak snack availability was in between these two time frames. 

It follows that the best day of the week to get free food of any kind is on Thursday, while the best day to get a free meal is on Wednesday. The week during the semester with the most abundant selection of free food is the first week of the Fall 2 quarter. Finally, the prime location to find free food is in the Kresge building.

We answered these questions through use of histograms and density graphs using ggplot2 and RShiny. After our analysis we went on to model our data using logistic regression with the likelihood of free food as our outcome of interest. With only 270 data points in our sample, we kept our model simple using only categorical time of day and categorical location as are our predictors of interest. The results of our model are below and show largely inconclusive data due in large part to limited sample size. Nevertheless, we attempted to create a predictive model from our data set by dividing it into a training and testing set with an 80/20 split. Our final model yields accuracy results of 60% and sensitivity and specificity at 69% and 50%, respectively. These results are lower than we would hope, but under our framework for predictive modeling, we do not have the statistical power necessary for a strong model.
  

```{r} 

#logistic regression 

statistics <-masterdata %>% select(ID, TimeStart, foodynfactor) %>% mutate(preten=ifelse(TimeStart<as.numeric(hms("10:00:00")), 1, 0), tentotwelve=ifelse(TimeStart>=as.numeric(hms("10:00:00")) & TimeStart<as.numeric(hms("12:00:00")), 1, 0), twelvetotwo=ifelse(TimeStart>=as.numeric(hms("12:00:00")) & TimeStart<as.numeric(hms("14:00:00")), 1, 0), twotofour=ifelse(TimeStart >= as.numeric(hms("14:00:00")) & TimeStart<as.numeric(hms("16:00:00")), 1, 0), fourtosix=ifelse(TimeStart>=as.numeric(hms("16:00:00")) & TimeStart<as.numeric(hms("18:00:00")), 1, 0), sixtoeight=ifelse(TimeStart>=as.numeric(hms("18:00:00")) & TimeStart<as.numeric(hms("20:00:00")), 1, 0), aftereight=ifelse(TimeStart >= as.numeric(hms("20:00:00")), 1, 0)) 

  
x<-glm(foodynfactor ~ tentotwelve+twelvetotwo+twotofour+fourtosix+sixtoeight+aftereight, data=statistics, family="binomial") 

x%>%summary 

  

exp(coef(x)) 

exp(confint(x)) 

``` 


```{r} 

## new dataframe for logistic analysis 

food_events1 <- event_df %>% select(-snack, -meal, -foodyn, -Leadership, -ODI) 

multianalysis <- left_join(statistics, food_events1, by="ID") 
multianalysis
food_events1

y<-glm(foodynfactor ~ tentotwelve+twelvetotwo+twotofour+fourtosix+sixtoeight+aftereight+Kresge+Countway+FXB+PHOffcampus, data=multianalysis, family="binomial") 

y%>%summary 


exp(coef(y)) 

exp(confint(y)) 

``` 


```{r} 

#Prediction with linear regression, training set and test set (JUST TIME) 

 set.seed(0) 

withfood<-statistics %>% filter(foodynfactor==1) %>% select(ID) %>% t() 

# 138 with food 

withoutfood<-statistics %>% filter(foodynfactor==0) %>% select(ID) %>% t() 

# 132 without food 

  

#sample 80% of events with food and 80% of events without food to make training dataset  

testfood<-sample(withfood, 106, replace=FALSE) 

testnofood<-sample(withoutfood, 106, replace=FALSE) 

trainset <- statistics %>% filter (ID %in% testfood | ID %in% testnofood)

testset <- statistics %>% filter (!(ID %in% testfood) & !(ID %in% testnofood)) 

  
## train set 

x<-glm(foodynfactor ~ tentotwelve+twelvetotwo+twotofour+fourtosix+sixtoeight+aftereight, data=trainset, family="binomial") 

x%>%summary 

library(caret) 

p_hat_logit <- predict(x, newdata = testset, type="response") 

y_hat_logit <- ifelse(p_hat_logit > 0.5, 1, 0) 

confusionMatrix(data = as.factor(y_hat_logit), reference = testset$foodynfactor, positive="1") 


``` 


```{r} 

trainset2 <- multianalysis %>% filter (ID %in% testfood | ID %in% testnofood) 

testset2 <- multianalysis %>% filter (!(ID %in% testfood) & !(ID %in% testnofood)) 

## train set 

z<-glm(foodynfactor ~ tentotwelve+twelvetotwo+twotofour+fourtosix+sixtoeight+aftereight+Kresge+Countway+FXB+PHOffcampus, data=trainset2, family="binomial") 

z%>%summary 

p_hat_logit2 <- predict(z, newdata = testset2, type="response") 

y_hat_logit2 <- ifelse(p_hat_logit2 > 0.5, 1, 0) 

confusionMatrix(data = as.factor(y_hat_logit2), reference = testset2$foodynfactor, positive="1") 

 ```
